Study 1 Analysis

### 2025-01-04 manuscript/appendix.qmd
add demo columns to s1

# widen by appliance, with separate row for value, family, and stage_avg

@@

    ‘
s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) 
s2s_clust <- s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) %>% .$cluster |> as.factor() 
s2s <- cbind(s2s,c=s2s_clust)

>>>>

```{r}
#| eval: false
# widen by appliance, with separate row for value, family, and stage_avg

@@

s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) 
s2s_clust <- s2s |>  ungroup() |> select(-id,-block,-refClass,-pct,-plan) |> scale(center = TRUE, scale = TRUE) |> kmeans(centers = 3) %>% .$cluster |> as.factor() 
s2s <- cbind(s2s,c=s2s_clust)

### 2025-01-04 nb/s1_survey.qmd
add demo columns to s1

s1 <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))
s1 <- s1 |> 
  mutate(
    state1 = case_when(
      Colorado == 1 ~ "Colorado",
      Massachusetts == 1 ~ "Massachusetts",
      California == 1 ~ "California",
      Texas == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      Colorado == 2 ~ "Colorado",
      Massachusetts == 2 ~ "Massachusetts",
      California == 2 ~ "California",
      Texas == 2 ~ "Texas",
      TRUE ~ NA_character_
    )) |> 
  mutate(refClass = case_when(
  ReferenceClass == 1 ~ "USD",
  ReferenceClass == 2 ~ "kWh",
  ReferenceClass == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  rename(id2=ResponseId) |>
  mutate(id=row_number()) |> relocate(id) |> 
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc, .after = id)

>>>>

s1 <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))
s1 <- s1 |> 
  mutate(
    state1 = case_when(
      Colorado == 1 ~ "Colorado",
      Massachusetts == 1 ~ "Massachusetts",
      California == 1 ~ "California",
      Texas == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      Colorado == 2 ~ "Colorado",
      Massachusetts == 2 ~ "Massachusetts",
      California == 2 ~ "California",
      Texas == 2 ~ "Texas",
      TRUE ~ NA_character_
    )) |> 
  mutate(refClass = case_when(
  ReferenceClass == 1 ~ "USD",
  ReferenceClass == 2 ~ "kWh",
  ReferenceClass == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01, 
         id2=ResponseId,
         gender=DEM01) |>
  mutate(id=row_number()) |> 
  relocate(id) |> 
  mutate(gender = factor(gender, 
                        levels = c(1,2,3),
                        labels = c("Male", "Female", "Not specified"))) |> 
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc,gender, .after = id)

### 2025-01-04 nb/study1.qmd
add demo columns to s1

s1 <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav"))

@@

s1 <- s1 |> 
  mutate(
    state1 = case_when(
      Colorado == 1 ~ "Colorado",
      Massachusetts == 1 ~ "Massachusetts",
      California == 1 ~ "California",
      Texas == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      Colorado == 2 ~ "Colorado",
      Massachusetts == 2 ~ "Massachusetts",
      California == 2 ~ "California",
      Texas == 2 ~ "Texas",
      TRUE ~ NA_character_
    )) |> 
  mutate(refClass = case_when(
  ReferenceClass == 1 ~ "USD",
  ReferenceClass == 2 ~ "kWh",
  ReferenceClass == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01) |>  # replace NA's in calc with 0's
  rename(id2=ResponseId) |>
  mutate(id=row_number()) |> relocate(id) |> 
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc, .after = id)

# Energy Literacy Score (ELS)

els <- s1 |> select(id,id2,calc,refClass,state1,state2, ELS01:ELS08)

@@

s1 <- s1 |> rename(edu=DEM04) |>
  mutate(edu = factor(case_when(
    edu == 1 ~ "Some schooling, but no diploma or degree",
    edu == 2 ~ "Highschool diploma or GED",
    edu == 3 ~ "Some college",
    edu == 4 ~ "College degree",
    edu == 5 ~ "Some graduate school",
    edu == 6 ~ "Graduate degree",
    edu == 7 ~ "Choose not to answer",
    TRUE ~ NA_character_
  ))) |> relocate(edu, .after = id)

@@

  df_long <- df %>%
  select(-matches("01$|02$|03$|04$|05$|06$")) %>%
    pivot_longer(
      cols = all_of(action_plan_cols),
      names_to = c("modality", "state", "item_num"),
      names_pattern = "^([ADP])\\d+(CAL|MASS|COL|TEX)(\\d+)$",
      values_to = "value"
    ) |> select(id,id2,state1,state2,refClass,calc,edu,modality,state,item_num,value,Duration__in_seconds_)

@@

#saveRDS(s1, here::here("data/s1_processed.rds"))

>>>>

s1 <- haven::read_sav(here::here("data/Frequency & Probability Study 1 - 3-24-19.sav")) |> select(-IPAddress)

@@

s1 <- s1 |> 
  mutate(
    state1 = case_when(
      Colorado == 1 ~ "Colorado",
      Massachusetts == 1 ~ "Massachusetts",
      California == 1 ~ "California",
      Texas == 1 ~ "Texas",
      TRUE ~ NA_character_
    ),
    state2 = case_when(
      Colorado == 2 ~ "Colorado",
      Massachusetts == 2 ~ "Massachusetts",
      California == 2 ~ "California",
      Texas == 2 ~ "Texas",
      TRUE ~ NA_character_
    )) |> 
  mutate(refClass = case_when(
  ReferenceClass == 1 ~ "USD",
  ReferenceClass == 2 ~ "kWh",
  ReferenceClass == 3 ~ "Percentage"
)) |> 
  rename(calc=MATH01, 
         id2=ResponseId,
         gender=DEM01,
         byear=DEM02,
         edu=DEM04,
         duration=Duration__in_seconds_) |>
  mutate(id=row_number(),
  age=as.numeric(format(RecordedDate, "%Y")) - byear) |> 
    mutate(edu = factor(case_when(
    edu == 1 ~ "Some schooling, but no diploma or degree",
    edu == 2 ~ "Highschool diploma or GED",
    edu == 3 ~ "Some college",
    edu == 4 ~ "College degree",
    edu == 5 ~ "Some graduate school",
    edu == 6 ~ "Graduate degree",
    edu == 7 ~ "Choose not to answer",
    TRUE ~ NA_character_
  ))) |>
  relocate(id) |> 
  mutate(gender = factor(gender, 
                        levels = c(1,2,3),
                        labels = c("Male", "Female", "Not specified"))) |> 
  mutate(calc=factor(ifelse(is.na(calc),0,calc))) |> 
  relocate(refClass,state1,state2,calc,gender,byear,age,RecordedDate,edu, .after = id)

# Energy Literacy Score (ELS)
els <- s1 |> select(id,id2,calc,refClass,state1,state2, ELS01:ELS08)

@@

@@

  df_long <- df %>%
  select(-matches("01$|02$|03$|04$|05$|06$")) %>%
    pivot_longer(
      cols = all_of(action_plan_cols),
      names_to = c("modality", "state", "item_num"),
      names_pattern = "^([ADP])\\d+(CAL|MASS|COL|TEX)(\\d+)$",
      values_to = "value"
    ) |> select(id,id2,state1,state2,refClass,calc,age,gender,edu,modality,state,item_num,value,duration)

@@

saveRDS(s1, here::here("data/s1_processed.rds"))
